<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{A Markdown Vignette with knitr}
-->

```{r setup, include=FALSE}
# set global chunk options
library(knitr)
opts_chunk$set(cache=FALSE)

## Function for embedding high qual text images:
uri_embed <- function(path, add="") {
    uri <- knitr::image_uri(path)
    cat(paste0("<img ", add, " src=\"", uri, "\" />"))
}
```

# Introduction to pacman
## Tyler Rinker & Dason Kurkiewicz

The **pacman** package is an R package management tool that combines the functionality of base `library` related functions into intuitively named functions.  This package is ideally added to [.Rprofile](http://stat.ethz.ch/R-manual/R-patched/library/base/html/Startup.html) to increase workflow by reducing time recalling obscurely named functions, reducing code, and integrating functionality of base functions to simultaneously perform multiple actions.  


```{r, echo=FALSE, results='asis'}
uri_embed("r_pacman.png", 
    "width=\"350\", height=\"150\" style=\"display:block; margin-left:auto; margin-right:auto;\"")
```

Function names in the pacman package follow the format of `p_xxx` where 'xxx' is the task the function performs.  For instance the `p_load` function allows the user to load one or more packages as a more generic substitute for the `library` or `require` functions and if the package isn't available locally it will install it for you.

## Installing, Loading, Unloading, Updating, & Deleting   

-Intro here- 

**Quick Reference Table** -Need to add better table styling-

| pacman Function | Base Equivalent | Description |
|----------------------|----------------------|----------------|
| `p_load`  |  `install.packages` + `library` | Loads and Install Packages |
| `p_install` |  `install.packages` | Install Packages from CRAN |
| `p_temp` |  NONE | Install a Package Temporarily |
| `p_unload` | `detach` | Unloads Packages from the Search Path |
| `p_update` | `update.packages` | Update Out-of-Date Packages |



### Installing and Loading    

- `p_load` 
- `p_install`/`p_get`
- `p_temp`
       
###  Unloading*

- `p_unload`

### Updating    
  
- `p_up`/`p_update`

### Deleting   
  
- `p_del`/`p_delete`

## Session Information    

- ` p_loaded`

##  Local Package Information    

-  `p_author`
-  `p_cite`/`p_citation`
-  `p_data`
-  `p_exists`
-  `p_funs`/`p_functions`
-  `p_help`
-  `p_info`/`p_information`
-  `p_inter`/`p_interactive`
-  `p_isloaded`
-  `p_news`
-  `p_ver`/`p_version`
-  `p_vign`/`p_vignette`


## Internet Based Package Information  

-  `p_cran`
-  `p_search_any`/`p_sa`

## Library Information     

- `p_base`
- `p_lib`/`p_library`
- `p_path`
- `p_sl`/`p_search_library`
- `p_unlock`

## pacman Tools   

- `p_detectOS`    
- `p_isglobal`    
- `p_opendir`   


## References
```{r, echo=FALSE, message=FALSE}
#write.bibtex(file="references.bib")
```

```{r clean-up, include=FALSE}
# R compiles all vignettes in the same session, which can be bad
rm(list = ls(all = TRUE))
```
