<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{A Markdown Vignette with knitr}
-->

```{r setup, include=FALSE}
# set global chunk options
library(knitr)
opts_chunk$set(cache=FALSE)

## Function for embedding high qual text images:
uri_embed <- function(path, add="") {
    uri <- knitr::image_uri(path)
    cat(paste0("<img ", add, " src=\"", uri, "\" />"))
}
```

# Introduction to pacman
## Tyler Rinker & Dason Kurkiewicz

The **pacman** package is an R package management tool that combines the functionality of base `library` related functions into intuitively named functions.  This package is ideally added to [.Rprofile](http://stat.ethz.ch/R-manual/R-patched/library/base/html/Startup.html) to increase workflow by reducing time recalling obscurely named functions, reducing code, and integrating functionality of base functions to simultaneously perform multiple actions.  


```{r, echo=FALSE, results='asis'}
uri_embed("r_pacman.png", 
    "width=\"350\", height=\"150\" style=\"display:block; margin-left:auto; margin-right:auto;\"")
```

Function names in the pacman package follow the format of `p_xxx` where 'xxx' is the task the function performs.  For instance the `p_load` function allows the user to load one or more packages as a more generic substitute for the `library` or `require` functions and if the package isn't available locally it will install it for you.




## References
```{r, echo=FALSE, message=FALSE}
#write.bibtex(file="references.bib")
```

```{r clean-up, include=FALSE}
# R compiles all vignettes in the same session, which can be bad
rm(list = ls(all = TRUE))
```
