<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{A Markdown Vignette with knitr}
-->

```{r setup, include=FALSE}
# set global chunk options
library(knitr)
opts_chunk$set(cache=FALSE)

## Function for embedding high qual text images:
uri_embed <- function(path, add="") {
    uri <- knitr::image_uri(path)
    cat(paste0("<img ", add, " src=\"", uri, "\" />"))
}
opts_knit$set(upload.fun = image_uri, self.contained=TRUE)
```

<h1 class="title">Introduction to pacman</h1>
<h1 class="auth">Tyler Rinker & Dason Kurkiewicz</h1>

The **pacman** package is an R package management tool that combines the functionality of base `library` related functions into intuitively named functions.  This package is ideally added to [.Rprofile](http://stat.ethz.ch/R-manual/R-patched/library/base/html/Startup.html) to increase workflow by reducing time recalling obscurely named functions, reducing code, and integrating functionality of base functions to simultaneously perform multiple actions.  The function `p_load` is particulalry well suited for help formums and posts, as it will load and, if necessary, install missing packages.


```{r, echo=FALSE, results='asis'}
uri_embed("r_pacman.png", 
    "width=\"350\", height=\"150\" style=\"display:block; margin-left:auto; margin-right:auto;\"")
```


**Quick Reference Table** 

| pacman Function | Base Equivalent | Description |
|----------------------|----------------------|----------------|
| `p_load`  |  `install.packages` + `library` | Loads and Install Packages |
| `p_install` |  `install.packages` | Install Packages from CRAN |
| `p_temp` |  NONE | Install a Package Temporarily |
| `p_unload` | `detach` | Unloads Packages from the Search Path |
| `p_update` | `update.packages` | Update Out-of-Date Packages |


Function names in the pacman package follow the format of `p_xxx` where 'xxx' is the task the function performs.  For instance the `p_load` function allows the user to load one or more packages as a more generic substitute for the `library` or `require` functions and if the package isn't available locally it will install it for you.

## Installing, Loading, Unloading, Updating, & Deleting   

The heart of **pacman** is it's ability to reduce typing in package management actions.  The functions in this section greatly reduce typing and increase code readability.  For example, many blog posts begin coding with this sort of package call:

```{r eval=FALSE}
packs <- c("XML", "devtools", "RCurl", "fakePackage", "SPSSemulate")
success <- suppressWarnings(sapply(packs, require, character.only = TRUE))
install.packages(names(success)[!success])
sapply(names(success)[!success], require, character.only = TRUE)
```

With pacman this call can be reduced to:

```{r eval=FALSE}
library(pacman)
p_load(XML, devtools, RCurl, fakePackage, SPSSemulate)
```



### Installing and Loading    

- `p_load` 
- `p_install`/`p_get`
- `p_temp`
       
###  Unloading*

- `p_unload`

### Updating    
  
- `p_up`/`p_update`

### Deleting   
  
- `p_del`/`p_delete`

## Session Information    

- ` p_loaded`

##  Local Package Information    

-  `p_author`
-  `p_cite`/`p_citation`
-  `p_data`
-  `p_exists`
-  `p_funs`/`p_functions`
-  `p_help`
-  `p_info`/`p_information`
-  `p_inter`/`p_interactive`
-  `p_isloaded`
-  `p_news`
-  `p_ver`/`p_version`
-  `p_vign`/`p_vignette`


## Internet Based Package Information  

-  `p_cran`
-  `p_search_any`/`p_sa`

## Library Information     

- `p_base`
- `p_lib`/`p_library`
- `p_path`
- `p_sl`/`p_search_library`
- `p_unlock`

## pacman Tools   

- `p_detectOS`    
- `p_isglobal`    
- `p_opendir`   

```{r css, echo = FALSE}
options(markdown.HTML.stylesheet = "css/style.css")
```

## References
```{r, echo=FALSE, message=FALSE}
#write.bibtex(file="references.bib")
```

```{r clean-up, include=FALSE}
# R compiles all vignettes in the same session, which can be bad
rm(list = ls(all = TRUE))
```
